Attributions for Poverty in Post-Socialist Countries

### 2014-02-13 article2013.Rmd
Kirjoitin kappaleen elinoloista teoriaan ja lueskelin paperia eteenpäin. Meni aamua (9-15) sitten toisen projektin kimpussa, niin huomiselle siirtyi tämän valmistelua.
Pyrin huomenna kirjoittamaan menetelmien ja muuttujien osalta täsmennyksiä ja avaamaan analyysien tuloksia.

----

*Eriarvoisuus lähti tästä*

*Maaryhmittely on tästä pois*

**Tähän tulisi HDI:n/elinolojen perustelut**

----

Besides the so called regime-effect, national-level economic conditions have been proved to be associated with welfare attitudes. This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence. In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crises in distinct ways. While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong. Hence, by analyzing the role of changes in economic performance during the period of global financial crises, the research design of this article provides empirical evidence on the effects of crises to public perceptions of the causes of poverty. Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007). In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved. This result is confirmed by Kallio and Niemelä (2013). They found that the larger the economic growth in a country is, more likely people are to blame the poor themselves. Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping.

@@

Human Development Index           Measures life expectancy, education, and income indices as a composite statistic
UNDP

@@

@@

--------

*A more throughout description of modelling will be added as the final details in modelling will be decided*

@@

*Results and conclusions are not yet reshaped to match the recent changes in the analysis...*

@@

HERE WE NEED AN INTREPETATIIONS OF HDI-FIGURE. BUT THE RESOURCES FIRST

@@

>Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media. [info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf)

>>>>

Besides the so called good government, national-level economic conditions have been proved to be associated with welfare attitudes. This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence. In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crises in distinct ways. While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong. Hence, by analyzing the role of changes in economic performance during the period of global financial crises, the research design of this article provides empirical evidence on the effects of crises to public perceptions of the causes of poverty. Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007). In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved. This result is confirmed by Kallio and Niemelä (2013). They found that the larger the economic growth in a country is, more likely people are to blame the poor themselves. Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping.

In addition to macro level economic performance, the level overall of well-being and living conditions are found to be linked with poverty attributions. Lepianka et al. (2010:65,86) found out that people living in more developed countries they tend to attribute living in destitution rather to individual unluckiness than to modern progress, but they found no link between development and either type of blame explanations. In contrast to previous studies, variation in human development within our selection of countries if far greater. Therefore we are able to poor countries of Central Asia with well-off Slovenia and Czech Republic. Given the great differences in living conditions within the post-socialist space, we assume that in countries of low human development and fewer welfare institutions people tend to attribute poverty more with individual reasons than social. At the same time people experience of transition with welfare state retrenchment and increasing social and health problems may well have produced support for structural explanations, too.

`Joku toinen lähde olisi paikallaan tähän liittyen. Ja tuo ristiriitaisen hypoteesin voisi oikoa..`

@@

Human Development Index           Measures life expectancy, education, and income indices as a composite statistics                             UNDP

@@

`Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media. [info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf) - *leikkasin tulos-luvusta tän kappaleen*`

`From Lepianka 2010  Independent variables s.60`

`Direct measurements of national poverty levels for all 28 countries and relevant years were not readily found, and therefore we used the Human Development Index (HDI) as a proximal indicator. The HDI seeks to rank countries in terms of their level of development and combines normalized measures of life expectancy, educational attainment and GDP per capita for countries worldwide. There is evidence that the HDI is very strongly correlated to the more recently developed Human Poverty Index, and thus, to a large extent, measures the same construct (cf. Sahay et al., 2003: 90). Here, we use the average HDI score per country for the period 1994–1999. Data come from the World Bank’s World Development Reports 1994–1999; see www.worldbank.org.`

@@

@@

@@

`On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1). 6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.`

`From Lepianka 2010 Conclusions s. 86`

`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`

@@

### 2014-02-13 finalPlots.Rmd
Pientä päivtystä. Huomenna iso remmi päälle
Kirjoitin kappaleen elinoloista teoriaan ja lueskelin paperia eteenpäin. Meni aamua (9-15) sitten toisen projektin kimpussa, niin huomiselle siirtyi tämän valmistelua.
Pyrin huomenna kirjoittamaan menetelmien ja muuttujien osalta täsmennyksiä ja avaamaan analyysien tuloksia.

mapplot <- ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white", size=.3) +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic") +
  theme_minimal() + 
  theme(legend.position="top") +
  theme(legend.title=element_blank()) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  theme(legend.key.size = unit(3, "mm"))

pdf("figure/mapplot.pdf", width=14/2.54, height=12/2.54)
mapplot
dev.off()

>>>>

mapplot <- ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white", size=.3) +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,60)) +
  theme_minimal() + 
  theme(legend.position="top") +
  theme(legend.title=element_blank()) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  theme(legend.key.size = unit(3, "mm"))

pdf("figure/mapplot.pdf", width=12/2.54, height=7/2.54)
mapplot
dev.off()
